### 2019-02-26 Seneca.tex
chandras pass

To facilitate model search and selection using the serverless architecture, 
we have developed Seneca, 
a framework for tuning the hyperparameters of machine learning applications 
in AWS Lambda. 
The Seneca pipeline consists of packaging, deployment, 
function optimization, and hyperparameter tuning. Figure~\ref{fig:seneca} 
shows the architecture of Seneca.  In the the upper-right-front, we
show the three inputs that Seneca expects from its users: a hyperparameter
configuration file, a dataset URL, and the path to 
the machine learning application.
The configuration file
specifies the range of values for each hyperparameter that the 
application expects.  Seneca uses this configuration to construct
the hyperparameter search space.  The dataset URL refers to a
valid dataset stored in AWS S3.

@@

The Seneca optimizer first configures and invokes the function using 
a user-defined payload.  From this run, Seneca obtains the memory 
used by function (reported by AWS Cloudwatch), 
which it uses as its starting point.
Seneca then defines two double-ended queues 
(deque) of length N, to store 
allocated memory and compute charge data of different
invocations. While the current 
allocated memory is less than or equal to 3008 MB, 
the optimizer reconfigures and invokes the function 
using the next increment for memory allocation.  It 
calculates the compute charge for each invocation 
using current allocated memory and billed duration.

We employ two exit conditions.  The first is when the 
compute charge monotonically increases across the deque. 
The second is when the increase in slope is greater than a threshold. 
Once the optimizer finds that both conditions hold, 
it pops the left-most value from deque and configures 
the function to use that value for allocated memory for all future invocations.
After extensive experimentation, we find that 
N=5 and a slope threshold of 1 work well in most cases.
These values can be changed by users of the system however.
In addition, this optimization can be turned on or off via a command line
argument to Seneca.

\begin{algorithm}[]
\SetAlgoLined
\KwData{Typical payload}
\KwResult{Optimal allocated memory}
Find memory used by payload as start point\;
Define deques for allocated memory \& compute charge\;
 \While{allocated memory $\leq$ 3008 MB}{

  \eIf{compute charge monotonically increase in deque \& slope $\geq$ 1}{
   Popleft from deques\;
   Configure allocated memory of function\; 
   Exit\;
   }{
   Increase allocated memory by 64 MB\;
   Probe lambda function\;
   Append allocated memory and compute charge to deques\;
  }
 }
 \caption{Seneca Optimizer Heuristic}
 \label{algo:optimizer}
\end{algorithm}

>>>>

To facilitate model search and selection using the serverless architecture, 
we have developed Seneca, 
a framework for tuning the hyperparameters of machine learning applications 
in AWS Lambda. 
The Seneca pipeline consists of packaging, deployment, 
function optimization, and hyperparameter tuning. Figure~\ref{fig:seneca} 
shows the architecture of Seneca.  In the the upper-right-front, we
show the three inputs that Seneca expects from its users: a hyperparameter
configuration file, a dataset URL, and the path to 
the machine learning application.
The configuration file
specifies the range of values for each hyperparameter that the 
application expects.  Seneca uses this configuration to construct
the hyperparameter search space.  The dataset URL refers to a
valid dataset stored in the AWS Simple Storage 
Service (S3)\footnote{\url{https://aws.amazon.com/s3/}}.

@@

The Seneca optimizer first configures and invokes the function using 
a user-defined payload.  From this run, Seneca obtains the memory 
used by function (reported by AWS Cloudwatch), 
which it uses as its starting point.
Seneca then defines two double-ended queues 
(\texttt{deque}) of length N, to store 
allocated memory and compute charge data of different
invocations. While the current 
allocated memory is less than or equal to 3008 MB, 
the optimizer reconfigures and invokes the function 
using the next increment for memory allocation.  It 
calculates the compute charge for each invocation 
using current allocated memory and billed duration.

We employ two exit conditions.  The first is when the 
compute charge monotonically increases across \texttt{deque}. 
The second is when the increase in slope is greater than a threshold. 
Once the optimizer finds that both conditions hold, 
it pops the left-most value from \texttt{deque} and configures 
the function to use that value for allocated memory for all future invocations.
After extensive experimentation, we find that 
N=5 and a slope threshold of 1 work well in most cases.
These values can be changed by users of the system however.
In addition, this optimization can be turned on or off via a command line
argument to Seneca.

\begin{algorithm}[]
\SetAlgoLined
\KwData{Typical payload}
\KwResult{Optimal allocated memory}
Find memory used by payload as starting point\;
Define deque for allocated memory \& compute charge\;
 \While{allocated memory $\leq$ 3008 MB}{

  \eIf{compute charge monotonically increases in deque \& slope $\geq$ 1}{
   Popleft from deque\;
   Configure allocated memory of function\; 
   Exit\;
   }{
   Increase allocated memory by 64 MB\;
   Probe lambda function\;
   Append allocated memory and compute charge to deque\;
  }
 }
 \caption{Seneca Optimizer Heuristic}
 \label{algo:optimizer}
\end{algorithm}

### 2019-02-26 conclusion.tex
chandras pass

We present a new framework, called Seneca, for simplifying and expediting
the training and testing of machine learning models
using low cost cloud services.
Specifically, Seneca leverages the AWS Lambda for autoscaled
(elastic) and parallel execution of hyperparameter search 
and selection for machine learning models.  
Users provide Seneca with the application code and libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  
%Seneca uses this 
%information to automatically construct models using parallel functions in AWS Lambda. 
%an AWS Lambda function that
%that invokes the application, imports and splits the dataset into training
%and testing datasets, and constructs, tests, and evaluates (i.e. scores) a machine
%learning model for a particular configuration of hyperparameter settings.  
Seneca uses this information to orchestrate parallel execution of these functions for all possible 
combinations of hyperparameter values specified and returns the best scoring model
and configuration to the user (for future use on other datasets).

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification (using
the models for prediction and not explanatory value), shows that Seneca is able to quickly
identify the best performing model for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

>>>>

We present a new framework, called Seneca, for simplifying and expediting
the training and testing of machine learning models
using low cost cloud services.
Specifically, Seneca leverages the AWS Lambda for autoscaled
(elastic) and parallel execution of hyperparameter search 
and selection for machine learning models.  
Users provide Seneca with the application code and 
libraries, 1+ datasets, and the list 
of possible hyperparameter settings.  
Seneca uses this information to automatically
configure and deploy these functions in parallel for all possible 
combinations of hyperparameter values specified. Seneca
returns the best scoring model
and configuration to the user (for future use on other datasets).

We present the design, implementation, and cost optimization for Seneca.
The optimizer automatically optimizes function memory use to reduce the cost
of AWS Lambda use.  Our empirical evaluation using
multiple applications for regression and classification,
shows that Seneca is able to quickly
identify the best performing hyperparameter setting 
for the applications and datasets that we consider.  
We also show that Seneca memory optimization reduces the cost of using
Seneca by 10-30\% for the applications studied.

### 2019-02-26 related_work.tex
mz - pass after 30 execution time
chandras pass

In this section, we overview related work. We discuss contributions that focus
on serverless computing and machine learning model optimization.

Multiple researchers have investigated the efficacy and overhead of the
serverless programming model and 
implementations~\cite{ref:jonas2017occupy,ref:onesteptwostep, ref:baldini2017,ref:lin2018tracking}.
The authors in \cite{ref:onesteptwostep} evaluate the cost and performance of using
serverless for data-intensive computing.  They identify some key weaknesses in AWS
Lambda for this application domain including the inability to effectively train models.
Our work show that it is possible for the subset of machine learning applications that we 
consider, to train models in a cost effective ways.

%\cite{ref:baldini2017} identifies multiple open research questions surrounding serverless,
%including whether or not functions should be stateless. 
PyWren~\cite{ref:jonas2017occupy} provides us a new perspective of stateless function being a natural fit for distributed computing models, because it 
abstracts away cluster management overhead and is ideal for embarrassingly parallel jobs. 
ExCamera~\cite{ref:encoding} presents a 
framework for running general-purpose parallel tasks (encoding 4K video) 
on a commercial serverless platform using multithreading. 
An empirical comparison of resource utilization and performance isolation
is provided in~\cite{ref:peeking} for AWS Lambda, 
Azure Functions, and Google Cloud Functions.
%to discuss the resource utilization and performance isolation efficiency of 
%three platforms. Besides efficacy and efficiency, there are challenges serverless architecture are facing in terms of 
%application monitoring and distributed storage. GammaRay~\cite{ref:lin2018tracking} proposes a cloud service that extracts
%causal dependencies across serverless functions and cloud services without programmer intervention. Such monitoring and 
%debugging tool simplifies the development and maintenance of serverless applications. Pocket~\cite{ref:ephemeral} provides 
%an elastic and economical distributed data store for severless analytic applications. We expect more contributions to toolchain 
%and distributed storage of serverless in upcoming years.

%please also discuss and cite these papers - you may need a second paragraph
%http://www.cs.ucsb.edu/~ckrintz/papers/ic2e18.pdf
%http://shivaram.org/publications/pywren-socc17.pdf
%S. Fouladi, R. S. Wahby, B. Shacklett, K. V. Balasubramaniam, W. Zeng, R. Bhalerao, A. Sivaraman, G. Porter, and K., Winstein. “Encoding, Fast and Slow: Low-Latency Video Pro- cessing Using Thousands of Tiny Threads”. In: NSDI. Boston, MA: USENIX Association, 2017, pp. 363–376.
%A. Klimovic, Y. Wang, C. Kozyrakis, P. Stuedi, J. Pfefferle, and A. Trivedi. “Understanding Ephemeral Storage for Serverless Analytics”. In: USENIX ATC. Boston, MA: USENIX Associa- tion, 2018, pp. 789–794.
%. Wang, M. Li, Y. Zhang, T. Ristenpart, and M. Swift. “Peek- ing Behind the Curtains of Serverless Platforms”. In: USENIX ATC. Boston, MA: USENIX Association, 2018, pp. 133–146.

The serverless framework~\cite{ref:serverless_framework} contributes an 
automated packaging and deployment toolset for serverless functions across clouds. 
The serverless framework uses Cloudformation~\cite{ref:cloudformation} to package and deploy Lambda function, which introduces additional cost. Seneca uses local docker container to avoid this cost
and overhead. The cloud infrastructure provisioning framework Terraform~\cite{ref:terraform} 
also provides automated deployment of functions to serverless platforms, but it does not simulate the AWS Lambda runtime to create the deployment package (doing so on a non-Linux system will result in Lambda errors and failures at execution time). 
In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead
to do so.

Automated hyperparameter optimization has been researched and developed 
in numerous projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
proposes a Python library for hyperparameter tuning. 
This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
which is a more advanced Python library for serial and parallel hyperparameter 
optimization over awkward search space. Hyperas~\cite{ref:hyperas} adds another 
abstraction layer to hyperopt and facilitates the hyperparameter tuning for 
Keras~\cite{ref:keras}. However, no extant work of which we are aware leverages
serverless to perform hyperparameter tuning in parallel.

>>>>

We next overview related work. We discuss contributions that focus
on serverless computing, automatic packaging, 
and machine learning (ML) model optimization.

Multiple researchers have investigated the efficacy and overhead of the
serverless programming model and 
implementations~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:baldini2017,ref:lin2018tracking}.
%The authors in \cite{ref:onesteptwostep} evaluate the cost and performance of using
%serverless for data-intensive computing.  
They identify some weaknesses in AWS
Lambda for data-intensive computing, including the inability to 
effectively train models.
Our work shows that it is possible and cost effective to use
AWS Lambda to train models for the subset of ML applications that we 
consider.

PyWren~\cite{ref:jonas2017occupy} shows that stateless functions
are a natural fit for distributed computing models. The technique
abstracts away cluster management overhead and is ideal for 
embarrassingly parallel jobs.  ExCamera~\cite{ref:encoding} presents a 
framework for running general-purpose parallel tasks (encoding 4K video) 
on a commercial serverless platform using multithreading. 
Cirrus~\cite{ref:jonas2019cloud} attempts to train ML models 
using a parameter server and serverless functions. 
An empirical comparison of resource utilization and performance isolation
is provided in~\cite{ref:peeking} for AWS Lambda, 
Azure Functions, and Google Cloud Functions.

The serverless framework~\cite{ref:serverless_framework} provides
automated packaging and deployment for serverless functions across clouds. 
The serverless framework uses CloudFormation\footnote{\url{https://aws.amazon.com/cloudformation/}} for deployment in AWS Lambda, which introduces additional cost. 
The cloud infrastructure provisioning 
framework Terraform\footnote{\url{https://www.terraform.io/}}
also provides automated deployment of functions to serverless platforms
(however, without replicating the Lambda runtim, introducing deployment errors).
Seneca uses a local docker container to avoid this cost and overhead. 
%In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead to do so.

Automated hyperparameter optimization is the focus of
many projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
and Hyperopt\footnote{\url{http://hyperopt.github.io/hyperopt/}}
provide a Python library for hyperparameter tuning. 
%This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
%which is a more advanced Python library for serial and parallel hyperparameter 
%optimization over awkward search space. 
Hyperas\footnote{\url{http://maxpumperla.com/hyperas/}} adds another 
abstraction layer to hyperopt to facilitate hyperparameter tuning for 
Keras~\footnote{\url{https://keras.io/}}. However, no extant work of which we are aware leverages
serverless to perform hyperparameter tuning and memory optimization in parallel 
for ML applications.
